MS Keyrun - GB survey footprint

### 2020-11-13 docs/GBSurveySet.Rmd
added plot

### Option 1 - use survey design
The NEFSC bottom trawl survey uses a stratified random design.  The rationale behind a stratified random design is to minimize variance. Figure \@ref(fig:Survey_map) shows the full range of strata with the Georges Bank EPU overlaid in red. As you will notice, some of the survey strata extend beyond the EPU boundary.  There are several ways that this can be accounted for.  The simplest option is to include all strata that are partially within the EPU boundary.

@@

Another option is to include only those strata that are at least 50% within the EPU boundary by area.  This was the preferred method of the Northeast IEA program for their State of the Ecosystem report.  The map below shows the strata that were included (Figure \@ref{fig:SOE}).  The medium blue color represents Georges Bank.

@@

>>>>

### Option 1 - use survey design
The NEFSC bottom trawl survey uses a stratified random design.  The rationale behind a stratified random design is to minimize variance. The figure below shows the full range of strata with the Georges Bank EPU overlaid in red. As you will notice, some of the survey strata extend beyond the EPU boundary.  There are several ways that this can be accounted for.  The simplest option is to include all strata that are partially within the EPU boundary.

@@

Another option is to include only those strata that are at least 50% within the EPU boundary by area.  This was the preferred method of the Northeast IEA program for their State of the Ecosystem report.  The map below shows the strata that were included.  The medium blue color represents Georges Bank.

@@

## Testing the options
To see if there was an noticeable difference between the three options outlined above (All - Any strata that is at least partially within the GB EPU boundary, SOE - strata set used by the State of the Ecosystem report, and EPU - post-stratified EPU) biomass indices were calculated for the top 11 species that will be included in the MS Keyrun project.

```{r test}
library(data.table); library(here); library(rgdal); library(Survdat);
library(ggplot2)

#source(here('Survdat.r'))
load(here('data_raw/survdat.RData'))

#Top 10(11) species
top10 <- c(15, 23, 32, 72, 73, 74, 75, 105, 106, 121, 197)

#fix numbers as character
survdat <- survdat[, STRATUM := as.numeric(STRATUM)]
survdat <- survdat[, SVSPP   := as.numeric(SVSPP)]

#First with survey strata
strata <- readOGR(here('gis'), 'strata', verbose = F)

#Generate area table
strat.area <- getarea(strata, 'STRATA')
setnames(strat.area, 'STRATA', 'STRATUM')

#Run 1 - All strata for GB ()
GB.soe   <- c(1090, 1130:1210, 1230, 1250, 3460, 3480, 3490, 3520:3550)
GB.extra <- 1100:1120

#Subset by season/ strata set
fall.all <- survdat[SEASON == 'FALL' & STRATUM %in% c(GB.soe, GB.extra), ]

#Run stratification prep
all.prep <- stratprep(fall.all, strat.area, strat.col = 'STRATUM', area.col = 'Area')

#Calculate stratified means
all.mean <- stratmean(all.prep, groups = top10, group.col = 'SVSPP', 
                       strat.col = 'STRATUM')
all.mean <- all.mean[, list(YEAR, SVSPP, strat.biomass)]
all.mean[, subset := 'All']

#Run 2 - SOE strata for GB ()
#Subset by season/ strata set
fall.soe <- survdat[SEASON == 'FALL' & STRATUM %in% GB.soe, ]

#Run stratification prep
soe.prep <- stratprep(fall.soe, strat.area, strat.col = 'STRATUM', area.col = 'Area')

#Calculate stratified means
soe.mean <- stratmean(soe.prep, groups = top10, group.col = 'SVSPP', 
                      strat.col = 'STRATUM')
soe.mean <- soe.mean[, list(YEAR, SVSPP, strat.biomass)]
soe.mean[, subset := 'SOE']

#Run 3 - Use EPU strata
#Grab strata
epu <- readOGR(here('gis'), 'EPU_strata', verbose = F)

#Generate area table
epu.area <- getarea(epu, 'Strata')
setnames(epu.area, 'Strata', 'EPU')

#Post stratify data if necessary
survdat.epu <- poststrat(survdat, epu, strata.col = 'Strata')
setnames(survdat.epu, 'newstrata', 'EPU')

#Subset by season/ strata set
fall.epu <- survdat.epu[SEASON == 'FALL', ]

#Run stratification prep
epu.prep <- stratprep(fall.epu, epu.area, strat.col = 'EPU', area.col = 'Area')

#Calculate stratified means
epu.mean <- stratmean(epu.prep, groups = top10, group.col = 'SVSPP', 
                      strat.col = 'EPU', poststrat = T)
epu.mean <- epu.mean[, list(YEAR, SVSPP, strat.biomass)]
epu.mean[, subset := 'EPU']

#Combine data
combo.data <- rbindlist(list(all.mean, soe.mean, epu.mean))

survey.plot <- ggplot(combo.data,
                      aes(x = YEAR, y = strat.biomass)) +
    geom_point(aes(color = subset)) +
    facet_wrap(~SVSPP, ncol = 2, nrow = 6, scales = 'free_y')

plot(survey.plot)

As you can see, all three spatial footprints give similar trends with only minimal differences in the magnitudes.  Therefore it was decided at a working group meeting on 11/4/2020 to use the survey strata set from the State of the Ecosystem report.  This accomplishes two goals.  First it stays consistent with the survey design of the survey.  Second it is consistent with other ecosystem products coming out of the Center.
